### 2018-04-16 experiments.tex
nodnet

The {\tt datarray} module \cite{datarray} is a prototype
implementation for numpy arrays with named axes to improve data
management, developed by the Berkeley Institute for Data Science.  As part of its code, it provides a set of algorithms for
inference in
Bayesian belief networks \cite{russell2016artificial}.  An earlier
version of these algorithms produced nondeterministic (and in some
cases incorrect) results due to dependence on the order of values in
an iterator over a Python dictionary.  This is a simple, clean example
of a problem that appeared in numerous Python programs when Python
version 3.3 introduced automatic random salting of hashes per-process,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries\footnote{In version 3.6, the default dictionary
  implementation became an ordered dictionary with consistent
  iteration, though actual hashes remained nondeterministic.}.  While
relatively few programs relied on exact hash values, many implicitly
relied on them in that they only functioned correctly with a
predictable order for dictionary iteration across processess.

>>>>

The {\tt datarray} module \cite{datarray} is a prototype
implementation for numpy arrays with named axes to improve data
management, developed by the Berkeley Institute for Data Science.  As part of its code, it provides a set of algorithms for
inference in
Bayesian belief networks \cite{russell2016artificial}.  An earlier
version of these algorithms produced nondeterministic (and in some
cases incorrect) results due to dependence on the order of values in
an iterator over a Python dictionary (see Section \ref{sec:pnondet}).

### 2018-04-16 implementation.tex
closer to submittable
ok
nodnet

@@

>>>>

\subsection{Sources of Nondeterminism in Practice}

\subsubsection{Process-Based Nondeterminism}

\label{sec:pnondet}

Some sources of nondeterminism require executing a test in a new
process environment, because the source is inherently tied to the
process environment.

Address Space Layout Randomization (ASLR) \cite{ASLR}  is probably the most
important source of nondeterminism that arises (only) from change in
process.  ASLR scrambles the layout of

Different CPU.

This is a simple, clean example
of a problem that appeared in numerous Python programs when Python
version 3.3 introduced automatic random salting of hashes (see Section
\ref{sec:pnondet} per-process,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries\footnote{In version 3.6, the default dictionary
  implementation became an ordered dictionary with consistent
  iteration, though actual hashes remained nondeterministic.}.  While
relatively few programs relied on exact hash values, many implicitly
relied on them in that they only functioned correctly with a
predictable order for dictionary iteration across processess.

\subsection{Implementing Nondeterminism Detection}

@@

Because TSTL has an interface to afl \cite{aflfuzz}, we can even use
afl's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism, using strategies built to find subtle
security vulnerabilities in C programs.
